Sperm should evolve to make female meiosis fair.

### 2014-05-26 writeup_and_notes/male_dependent_drive.tex
fixing typos found by Jeremy

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
	creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations in
	likely precludes this interaction in plants). 
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad), 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for exploitation by such selfish systems.  
Why then is the requirement of fertilization to complete female meiosis so ubiquitous?

@@

It seems that allowing sperm to influence the outcome of female meiosis would generate a confluence of these potential conflicts -- 
	sperm could actually assist an allele that distorts female meiosis.
However, this is not the case.
We find that an allele which acts through sperm to distort female meiosis in its favor %YB [too awkward / unclear?] 
	can rarely spread through a population if it bears any cost. 
Additionally, when this self-promoting driver can spread, it can only rarely 
	be maintained as a protected polymorphism, and due to its positive frequency dependence,  
	it spends very little time at intermediate frequency.
As such, this type of exploitation cannot generate a sustained genetic conflict.
It is therefore unlikely
	that female oogenesis and meiosis will evolve to prevent their effect.  
This likely explains why females can delay the completion of meiosis until after fertilization 
	without risking exploitation by collaborations between female
        drivers and sperm alleles.
%Male modification of recombination rates. \yb{trying to think through this \dots{}}

@@

Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before fertilization is complete. 
We argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness.  In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions. 
We predict that female drive, at a stage influenceable by sperm,  
	should be reduced when fertilized by
	complimentary sperm. 
Therefore, for example, MII female drive should be exceptionally severe in
	interspecific crosses between animals which complete the second meiotic division after fertilization,  
	as heterospecific sperm will not specifically evolve to resists a female driver in another species. 
Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
For example, we predict that a lower opportunity for female meiotic drive,
        e.g. an animal lineage with a long history of high inbreeding
        or selfing, 
        would often be accompanied by a relaxation of the requirement of fertilization for the
        completion of female meiosis.

>>>>

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
	creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations
	likely precludes this interaction in plants). 
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad), 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for exploitation by such selfish systems.  
Why then is the requirement of fertilization to complete female meiosis so ubiquitous?

@@

It seems that allowing sperm to influence the outcome of female meiosis would generate a confluence of these potential conflicts -- 
	sperm could actually assist an allele that distorts female meiosis.
However, this is not the case.
We find that an allele which acts through sperm to distort female meiosis in its favor %YB [too awkward / unclear?] 
	can rarely spread through a population if it bears any cost. 
Additionally, when this self-promoting driver can spread, it can only rarely 
	be maintained as a protected polymorphism, and due to its positive frequency dependence,  
	it spends very little time at intermediate frequency.
As such, this type of exploitation cannot generate a sustained genetic conflict.
It is therefore unlikely
	that female o\"{o}genesis and meiosis will evolve to prevent their effect.  
This likely explains why females can delay the completion of meiosis until after fertilization 
	without risking exploitation by collaborations between female
        drivers and sperm alleles.
%Male modification of recombination rates. \yb{trying to think through this \dots{}}

@@

Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before meiosis is complete. 
We argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness.  In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions. 
We predict that female drive, at a stage influenceable by sperm,  
	should be reduced when fertilized by
	complimentary sperm. 
Therefore, for example, MII female drive should be exceptionally severe in
	interspecific crosses between animals which complete the second meiotic division after fertilization,  
	as heterospecific sperm will not specifically evolve to resists a female driver in another species. 
Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
For example, we predict that a lower opportunity for female meiotic drive,
        e.g. an animal lineage with a long history of high inbreeding
        or selfing, 
        would often be accompanied by a relaxation of the requirement of fertilization for the
        completion of female meiosis.
